### 2023-04-29 my_paper/Hadah-last-names.tex
edits
edit paper to match thesis

@@

\begin{abstract}
In this paper, I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name. While men born to Hispanic father-White mothers earn less than those born to White father-Hispanic mothers, the gap could be explained entirely by educational differences. I also study the effect of identifying as Hispanic on earnings. I find that men that identify as Hispanic earn significantly less than those that do not, even after controlling for educational differences. 
\end{abstract}

@@

In this paper, I answer the following questions. Does having a Hispanic last name affect labor market outcomes? What is the effect of identifying as Hispanic on earnings? Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics. In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics. They find gaps in education and health between Hispanics and Whites. They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health. \citet{davilaChangesRelativeEarnings2008}, the authors documents many gaps in labor market outcomes between Hispanics and Whites. They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.

@@

Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills. one strategy used by researchers is audit or resume studies. \citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black Sounding names. This approach, however, has its drawbacks. Audit studies only observe callbacks, not wages.

This study utilizes a method developed by \citet{rubinstein2014pride}. I compare children from inter-ethnic marriages. More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH). This approach stems from the fact that marriages are not random. Couples match on several observable characteristics like income, schooling, socio-economic background, etc. \citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}. Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogonous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers. Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.

Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970-1990 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes. In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings. I also study the effect of identifying as Hispanic on labor market outcomes.

@@

On the one hand, WW children are more educated, with approximately 14 years of education for men and women. WW children have an employment rate of 95\% for men and 96\% for women. They have log hourly earnings of 2.48 for men and 2.33 for women. Also, their annual log earnings are 10.25 for men and 9.46 for women.

On the other hand, children of HH marriages have an average of 12.1 and 13.27 years of education for men and women, respectively. HH men and women have an employment rate of 93\% and 94\%. Their annual log earnings are 10.01 and 9.53 for men and women, respectively.

@@

WH children have an average of 13.58 and 13.87 years of education for men and women. The employment rate for men is 94\% and for women 95.0\%. HW men and women obtained 13.21 and 13.42 years of schooling. Men's employment rate is 92\% and women's 93\%.

In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic. \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 97\%, identify as Hispanic. There is considerable attrition in Hispanic identification of inter-ethnic children. Among WH children, 81\% of men and 85\% of women identify as Hispanic. Among HW children, 88\% of men and 87\% of women identify as Hispanic. These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}. Finally, a small portion of WW children, 5\% of men and 6\% of women, identified as Hispanic. These 6\% of WW children are most likely third generation--- or higher--- children of immigrants. For my analysis, I dropped the WW observations identify as Hispanic to compare with non-Hispanic Whites.

@@

The differentials between WH and HW are not as severe as the ones between HH and WW. HW children are still worse off than WH but are much better than HH. HW men attain 0.36 fewer years of education than WH men. HW women achieve 0.46 fewer years of schooling than WH women. HW children are 2 percentage points less likely to be employed. HW men earn 4\% less annual earnings as WH men, and HW women earn 7\% less. These differences show that HW and WH children are different from children of endogamous/homogonous marriages and, thus, are more comparable.

@@

 %In columns five and six of table \ref{tab:c1}, I present the gaps between HH and WW children and HW and WH children, respectively. The gap between HH and WW children (column five) shows a consistent story, HH children are worse in education and work more. HH children have more kids. While the gap between HW and WH does not have a clear-cut direction. Even though WH children obtain more years of education and have a higher employment rate, HW and WH men have the same fertility rates. HW women have a higher fertility rate than WH women. WH and HW women have the same employment rate.

@@

Using the 1970 and 1990 censuses, I constructed a data set of synthetic parents. The sample includes married White men and women. Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity. I consider a Hispanic person as White and born in a Spanish-speaking country. Consequently, Whites in the sample are people who are White and native-born. Using the information provided in the census, I can link husbands and wives with each other. I assume that parents have children between the ages of 25 and 40. Therefore, my sample consists of married White men and women that were born in the 1930 to 1965 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.

I show the distribution of the four types of couples in table \ref{tab:mat1}. White husbands and White wives (WW) make up the majority of couples in the sample, 97\% (1,286,731 couples). Hispanic husbands and wives (HH) are the second largest group making up 2\% (20,911 couples)of all couples. White husbands and Hispanic wives (WH) couples are less than 1\% (7,178 couples) of the sample and Hispanic husbands and White wives (HW) are also less than 1\% (7,606 couples).

I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}'s parent's panel. WW couples have higher education, 13.05 years for husbands and 12.74 for wives. As a household, WW couples have 25.78 years of schooling. Husbands in HH marriages have 8.93 years of education, while women have 8.6 years of schooling. As a household, HH couples have 17.54 years of education.

@@

In this section, I present two empirical strategies. The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings. The second empirical strategy will estimate the effect of identifying as Hispanic on log annual earnings.

@@

In this section I restrict the sample to WH and HW groups. Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$. $HW_{ist}$ is an indicator variables for the type of parents person $i$ has. $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term. The equation for this strategy is written as follows:

\begin{equation} \label{eq:1a}
Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
\end{equation}

@@

The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$. The coefficients $\beta_{2}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names.

@@

In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}. The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}. All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are waged and salaried workers.

@@

I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}. The sample in this analysis includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE). I also restrict the sample to children of HW and WH (omitted) parents. Column one in table \ref{tab:lastnamereg} is the results without controlling for education. Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name. This gap, however, could be entirely explained by educational differences. An inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.

I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}. The sample in this analysis includes full-time, full-year, and waged and salaried men, and I control for hours worked and age and include year fixed effects (FE). I also restrict the sample to children of HW and WH (omitted) parents. Column one in table \ref{tab:identreg} is the results without controlling for education. Overall, an inter-ethnic that identifies as Hispanic earns 11 percentage points less than an inter-ethnic that does not. This gap, however, could be entirely explained by educational differences. An inter-ethnic that identifies as Hispanic earns 6 percentage points less than an inter-ethnic that does not, but the result is statistically insignificant.

@@

I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name. The earning gap between WH and WW is 14\%, 20\% between HW and WW, and 24\% between HH and WW. When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6\% less than WH children. Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6\% less than someone with a White last name. Controlling for education, however, explains more than half of the earnings gaps between WW children and WH, HW, and HH. When I control for education, the last name effect decreases to a statistically insignificant 2\% earnings gap.

>>>>

\usepackage{lscape}
\usepackage{threeparttablex}
\usepackage{longtable}

@@

\begin{abstract}
In this paper, I study discrimination against Hispanics in the labor market. I compare the children of inter-ethnic marriages to study the impact of having a Hispanic last name. While males born to Hispanic father-White mothers earn less than those born to White father-Hispanic mothers, the gap could be completely explained by educational differences. I also study the effect of identifying as Hispanic on earnings. I find that men who identify as Hispanic earn significantly less than those who do not, even after controlling for educational differences.
\end{abstract}

@@

In this paper, I answer the following questions. Does having a Hispanic last name affect labor market outcomes? What is the effect of identifying as Hispanic on earnings? Moreover, I aim to show that comparing Hispanic Whites to non-Hispanic Whites might create an artificially higher earnings gap since the two groups differ on many observable characteristics.\footnote{By identifying as Hispanic, I refer to individuals who self-report their ethnicity as Hispanic on surveys or other data collection instruments.} \footnote{Observable characteristics refer to factors that can be measured and quantified, such as education level, work experience, and immigration status.} Others have attempted to compare how native-born White Hispanics fare to non-Hispanic Whites and foreign-born Hispanics. In \citet{antman2020ethnic,antmanEthnicAttritionObserved2016,antmanEthnicAttritionObserved2016a,antmanEthnicAttritionAssimilation2020}, the authors compare the health and educational outcomes of Hispanic Whites to non-Hispanic Whites and native-born Hispanics to foreign-born Hispanics. They find gaps in education and health between Hispanics and Whites. They also find that native-born Hispanics are more likely than their foreign-born counterparts to report poor health. \citet{davilaChangesRelativeEarnings2008} documents many gaps in labor market outcomes between Hispanics and Whites. They attribute a big part of this gap to differences in education, experience, immigration status, and regional differences.

@@

Identifying discrimination is difficult because of factors that affect labor market outcomes that are unobservable to economists---such as unobserved skills. One strategy used by researchers is audit or resume studies. \citet{bertrand2004emily} conducted an audit study where identical resumes were sent to employers with White and Black-sounding names. This approach, however, has its drawbacks. Audit studies only observe callbacks, not wages.

This study utilizes a method developed by \citet{rubinstein2014pride}. I compare children from inter-ethnic marriages. More precisely, I compare children of Hispanic fathers and White mothers (henceforth HW) to children of White fathers and Hispanic mothers (henceforth  WH). This approach stems from the fact that marriages are not random. Couples match on several observable characteristics like income, schooling, socio-economic background, etc. \citep{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}. Children of HW and WH marriages have more similar observable characteristics than children of endogamous/homogamous marriages--- i.e., White fathers-White mothers and Hispanic fathers-Hispanic mothers. Moreover, children from a Hispanic father and White mother household will have a Hispanic last name from their fathers, enabling the investigation of how ethnic signals, such as having a Hispanic last name, affect annual log earnings.

Using the Current Population Survey (CPS) from 1994 to 2019 and the 1970 to 1990 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes. In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings. I also study the effect of identifying as Hispanic on labor market outcomes.

@@

On the one hand, WW children are more educated, with approximately 14 years of education for men and women. WW children have an unemployment rate of 5\% for men and 4\% for women. They have log hourly earnings of 2.48 for men and 2.33 for women. Also, their annual log earnings are 10.25 for men and 9.46 for women.

On the other hand, children of HH marriages have an average of 12.1 and 13.27 years of education for men and women, respectively. HH men and women have an unemployment rate of 7\% and 6\%. Their annual log earnings are 10.01 and 9.53 for men and women, respectively.

@@

WH children have an average of 13.58 and 13.87 years of education for men and women. The unemployment rate for men is 6\% and for women 5\%. HW men and women obtained 13.21 and 13.42 years of schooling. Men's unemployment rate is 8\% and women's 7\%.

In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic. \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 97\%, identify as Hispanic. There is considerable attrition in Hispanic identification of inter-ethnic children. Among WH children, 81\% of men and 85\% of women identify as Hispanic. Among HW children, 88\% of men and 87\% of women identify as Hispanic. These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \citep{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}. Finally, a small portion of WW children, 5\% of men and 6\% of women, identified as Hispanic. These 6\% of WW children are most likely third generation--- or higher--- children of immigrants. For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.

@@

The differentials between WH and HW are not as severe as the ones between HH and WW. HW children are still worse off than WH but are much better than HH. HW men attain 0.36 fewer years of education than WH men. HW women achieve 0.46 fewer years of schooling than WH women. HW children are 2 percentage points less likely to be employed. HW men earn 4\% less annual earnings as WH men, and HW women earn 7\% less. These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.

@@

@@

Using the 1970 to 1990 censuses, I constructed a data set of synthetic parents. The sample includes married White men and women. Even though the census asks a person whether they are Hispanic or not, I took advantage of the questions on place of birth to create a proxy for ethnicity. I consider a Hispanic person as White and born in a Spanish-speaking country. Consequently, Whites in the sample are people who are White and native-born. Using the information provided in the census, I can link husbands and wives with each other. I assume that parents have children between the ages of 25 and 40. Therefore, my sample consists of married White men and women that were born in the 1930 to 1965 cohorts\footnote{The construction of "synthetic parents" follows the method used by \citet{rubinstein2014pride}.}.

I show the distribution of the four types of couples in table \ref{tab:mat2}. White husbands and White wives (WW) make up the majority of couples in the sample, 97\% (1,286,731 couples). Hispanic husbands and wives (HH) are the second largest group making up 2\% (20,911 couples)of all couples. White husbands and Hispanic wives (WH) couples are less than 1\% (7,178 couples) of the sample and Hispanic husbands and White wives (HW) are also less than 1\% (7,606 couples).

I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}. WW couples have higher education, 13.05 years for husbands and 12.74 for wives. As a household, WW couples have 25.78 years of schooling. Husbands in HH marriages have 8.93 years of education, while women have 8.6 years of schooling. As a household, HH couples have 17.54 years of education.

@@

In this section, I present two empirical strategies. The first empirical strategy estimates the effect of having a Hispanic last name on annual log earnings. The second empirical strategy estimates the effect of identifying as Hispanic on log annual earnings.

@@

In this section I restrict the sample to WH and HW groups. Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$. $HW_{ist}$ is an indicator variables for the type of parents person $i$ has. $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term. The equation for this strategy is written as follows:
\begin{equation} \label{eq:1a}
Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
\end{equation}

@@

The coefficients of interest from equation \ref{eq:iden} are $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$. The coefficients $\beta_{1}$ and $\beta_{3}$ capture the effects of Spanish-sounding last names.

@@

In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}. The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}. All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.

@@

I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}. The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE). I also restrict the sample to children of HW and WH (omitted) parents. Column one in table \ref{tab:lastnamereg} is the results without controlling for education. Overall, an inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name. This gap, however, could be entirely explained by educational differences. An inter-ethnic with a Hispanic last name earns 6 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.

I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}. The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE). I also restrict the sample to children of HW and WH parents. The omitted group is children of WH parents that do not identify as Hispanic. Column one in table \ref{tab:identreg} is the results without controlling for education. Overall, an inter-ethnic that identifies as Hispanic earns 11 percentage points less than an inter-ethnic that does not. This gap, however, could be entirely explained by educational differences. An inter-ethnic that identifies as Hispanic---i.e. $HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result is statistically insignificant.

@@

I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name. When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children. Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name. When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap. Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name. I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic. The earnings gap, however, could be entirely explained by educational differences.
